\textbf{ \Large{Reducing data traffic in wireless structural health monitoring systems using embedded computing}}

### 2015-02-08 Paper/PaperJahrSchlichSHM.tex

\title{\textbf{ \Large{Decentralized autonomous sensor fault detection using neural networks}} }

@@

% 
%\sloppy
\setlength{\emergencystretch}{3pt}
\hyphenation{sample-rate auto-nomous}

@@

The dependability and the accuracy of structural health moni\-toring systems can be affected by sensor faults. 
In this paper, the design and implementation of a wireless structural health monitoring system, capable of decentralized autonomous fault detection, are presented. 
For self-detecting sensor faults, each sensor node predicts expected sensor data and compares it to the measured sensor data. 
The predictions are computed using neural networks based on measured sensor data of adjacent sensor nodes.
In laboratory experiments, devised to validate the proposed approach, several simulated sensor faults are detected.
These results indicate that the use of neural networks for fault detection increases the dependability and the accuracy of structural health monitoring systems.

@@

Over their lifetime, sensors can become inaccurate, faulty, or may even break.
A fault can be defined as a defect of a sensor that leads to an error. An error is the manifestation of a fault---an incorrect system state---that may result in a failure.
To ensure the dependability and the accuracy of the SHM system, sensor faults must be detected and isolated in real time. 
A well known approach to fault detection is the installation of physically redundant sensors.
Faulty sensors can be identified through the deviation of their measurements from the measurements of correlated sensors.
Physical redundancy, although efficient for sensor fault detection, causes increased installation and maintenance costs due to the multiple installation of sensors. 
Representing a more efficient approach, analytical redundancy typically uses mathematical functions mapping the characteristics of the structure and the correlations of the installed sensors. Virtual sensor measurements are computed for each sensor and then compared to the actual measurements. 
For example, finite element models can be used in combination with data from adjacent sensor nodes to calculate virtual measurements of a sensor
\citep{Smarsly2014}.

In this study, analytical redundancy is implemented into wireless sensor nodes based on artificial neural networks.
Artificial neural networks essentially consist of interconnected data processing units called neurons. 
The neurons are grouped in different layers; usually one input layer, a number of hidden layers, and one output layer.
Artificial neural networks are able to learn, which is achieved by adjusting the weights of the inter-neuron connections until a set of given input variables results in the desired output variables; for example, a neural network can be trained to approximate any mathematical functions with any level of accuracy \citep{Li2011}.

This paper is organized as follows:
First, a wireless structural health monitoring system is designed and implemented. 
Next, a neural network is implemented into each sensor node and trained to predict the sensor measurements of the specific node for detecting sensor faults in a decentralized manner. 
Then, the system is tested in laboratory experiments. 
Finally, the experimental results are discussed and future research directions are proposed.

@@

The SHM system is programmed object-oriented in Java. Object orientation uses objects, that are instanciated using classes as paradigm. A class includes methods, allowing the objects to perform actions, and attributes, specific data.
\autoref{fig:UML} describes the classes of the SHM system code.
The SHM system consists of two packages---\texttt{sensornode} and \texttt{basestation}.
A package organizes several Java classes that build a program.

The package \texttt{sensornode} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Communication} and \texttt{MainSpot}, which are embedded directly into the sensor nodes.
The \texttt{AccelerationSampler} class is responsible for measuring the acceleration.
There are two phases: \colorbox{cyan}{At} first, the acceleration is measured with a low sampling rate.
Once the acceleration exceeds a threshold, the \colorbox{cyan}{second phase is entered} by increasing the sampling rate. 
The measured values are stored into an array.
The different phases are indicatad by lighting different LEDs.
The \texttt{FFT} class performs a fast Fourier transform on the measured accelerations. 
With the transformed data, the magnitudes and the correlating frequencies of the measured oscillation are calculated.
Finally, the natural frequency \colorbox{cyan}{is} determined by extracting the maximal magnitude.
The \texttt{Communication} class opens a radio connection between the sensor node and the basestation to transfer data from the sensor node to the basestation.
For starting the operation of \colorbox{cyan}{the sensor node}, the entry point of the programm is the \texttt{startApp()} method in the \texttt{MainSpot} class. 
Within the \texttt{MainSpot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class and the \texttt{Communication} class are created to perform the measurement.

The package \texttt{basestation} runs on the host computer and operates the basestation.
It consists of the classes \texttt{Database\-Handler} and \texttt{MainBase}.
The \texttt{Database\-Handler} class establishes a connection to a MySQL database, creates a database table, if none with the specified name is available, and inserts data into the database table.
The entry point of the program is the \texttt{run()} method in the \texttt{MainBase} class. The \texttt{MainBase} class opens a radio connection between the basestation and the sensor nodes, receives data sent by the sensor nodes and creates an instance of \texttt{Database\-Handler} to insert the data into the database.

@@

\section*{Implementation and training of neural networks}

Neural networks are implemented into the sensor nodes using SNIPE\footnote{}, a open source Java Library.
\begin{figure}[ht]
    \centering
    \includegraphics{figures/neuralnetwork.pdf}
    \caption{Schematic drawing of an artificial neural network with three layers}
    \label{fig:neuralnetwork}
\end{figure}
\todo[inline]{Grafik wird an unser neuronales Netz angepasst}

%\begin{itemize}
%\item neurons
%\item layer
%\item weights
%\item activation function / identity function
%\end{itemize}

%\textbf{2. paragraph: proposed nn}
%
%\begin{itemize}
%\item SNIPE? implementation in java?
%\item activation function / identity function
%\end{itemize}
%
%\textbf{3. paragraph: learning general}
%
%\begin{itemize}
%\item set with test values
%\item iterative weight adjustment
%\end{itemize}
%
%\textbf{4. paragraph?: learning, proposed}

\newpage
\section*{Laboratory experiments}

@@

@@

This paper has shown a decentralized autonomous sensor fault detection strategy for structural health monitoring systems based on neural networks. 
Autonomous sensor fault detection has been realized by implementing a neural network into each sensor node.
The neural networks have been trained to predict expected sensor measurements to be compared to actual measurments, in order to detect sensor faults.
The SHM system has been verified with laboratory experiments, proving that sensor fault detection using neural networks can improve the dependability and the accuracy of structural health monitoring systems.

>>>>

\title{\textbf{ \Large{Reducing data traffic in wireless structural health monitoring systems using embedded computing}}}

@@

% 
%\sloppy
\setlength{\emergencystretch}{3pt}
\hyphenation{sample-rate auto-nomous moni-toring}

@@

Structural health monitoring systems
shm systems produce a lot of data
we reduce the data with fft
The dependability and the accuracy of structural health monitoring systems can be affected by sensor faults. 
In this paper, the design and implementation of a wireless structural health monitoring system, capable of decentralized autonomous fault detection, are presented. 
For self-detecting sensor faults, each sensor node predicts expected sensor data and compares it to the measured sensor data. 
In laboratory experiments, devised to validate the proposed approach, the data traffic is reduced.

@@

\begin{itemize}
	\item problem: viel data traffic
	\item daten können verloren gehen
	\item lösung: embedded computing, decentralized data analysis on sensor nodes 
	\item fft
\end{itemize}

This paper is organized as follows:
First, a wireless structural health monitoring system is designed and implemented. 
Next, a decentralized data analysis approach, using fast Fourier transform, is developed.
Then, the system is tested in laboratory experiments.
Finally, the experimental results are discussed and future research directions are proposed.

@@

The SHM system is programmed object-oriented in Java. 
Object orientation uses objects which are instanciated using classes as paradigm. 
A class includes methods, allowing the objects to perform actions, and attributes, storing object-specific data.
The SHM system consists of two packages---\texttt{sensornode} and \texttt{basestation}.
A package organizes several Java classes that build a program.

\autoref{fig:UML} describes the classes of the \texttt{sensornode} package.
The package \texttt{sensornode} consists of the classes \texttt{Acceleration\-Sampler}, \texttt{FFT}, \texttt{Communication} and \texttt{MainSpot}, which are embedded directly into the sensor nodes.
The \texttt{AccelerationSampler} class is responsible for measuring the acceleration.
There are two phases: \colorbox{cyan}{At} first, the acceleration is measured with a low sampling rate.
Once the acceleration exceeds a threshold, the \colorbox{cyan}{second phase is entered} by increasing the sampling rate. 
The measured values are stored into an array.
The different phases are indicatad by lighting different LEDs.
The \texttt{FFT} class performs a fast Fourier transform on the measured accelerations. 
With the transformed data, the magnitudes and the correlating frequencies of the measured oscillation are calculated.
Finally, the natural frequency \colorbox{cyan}{is} determined by extracting the maximal magnitude.
The \texttt{Communication} class opens a radio connection between the sensor node and the basestation to transfer data from the sensor node to the basestation.
For starting the operation of \colorbox{cyan}{the sensor node}, the entry point of the programm is the \texttt{startApp()} method in the \texttt{MainSpot} class. 
Within the \texttt{MainSpot} class, instances of the \texttt{Acceleration\-Sampler} class, the \texttt{FFT} class and the \texttt{Communication} class are created to perform the measurement.
\todo{Grafik zweitteilen - hier package sensornode}

\autoref{fig:UML} describes the classes of the \texttt{basestation} package.
The package \texttt{basestation} runs on the host computer and operates the basestation.
It consists of the classes \texttt{Database\-Handler} and \texttt{MainBase}.
The \texttt{Database\-Handler} class establishes a connection to a MySQL database, creates a database table, if none with the specified name is available, and inserts data into the database table.
The entry point of the program is the \texttt{run()} method in the \texttt{MainBase} class. The \texttt{MainBase} class opens a radio connection between the basestation and the sensor nodes, receives data sent by the sensor nodes and creates an instance of \texttt{Database\-Handler} to insert the data into the database.
\todo{Grafik zweitteilen - hier package basestation}

@@

\section*{Laboratory experiments}

In the following section, the laboratory experiments are described.
First a describtion of the test setup is given, second, the data aquisition and processing are depicted, and finally, the results are discussed.

@@

\begin{itemize}
	\item 
	\item that means our paper is also wireless and very very good
\end{itemize}

@@

\begin{itemize}
	\item data traffic was reduced by XX\,\% 
\end{itemize}
